\section*{Acknowledgements}
I am immensely thankful to my supervisors Dave Meger and Marc Bellemare for
their perpetual mentorship and guidance over the course of my Masters'. Over the
past year and a half of thesis work, I have repeatedly been impressed by
the patience they had for the many rabbit holes I got lost in, the lulls in
progress, and the occasional technical issues I've had due to my stubborn
computer preferences. I truly commend them for not giving up on me in the many
moments where I felt like my efforts were futile.

Thanks as well to Dave Meger, Greg Dudek, and Hsiu-Chin Lin for making the
Mobile Robotics Lab at McGill such an amazing environment to be a part of, as
well as the many MRL students that have always made me feel welcome. Special
thanks to Jean-François Tremblay, Wesley Chung, Melissa Mozifian, Sahand Rezaei
Shostari, Lucas Berry, Andrew Holliday, and Abhisek Konar for the many sessions
of discussion and guidance along the way. I hope to see you all in person one
day.

Marc Bellemare's groups at Mila and Google have also
been a great source of inspiration. I am very thankful for the several
discussions with Ross Goroshin about fancy techniques in continuous-time optimal
control, which have contributed a lot to my appreciation of the field. 

Additionally, I must thank Professors Prakash Panangaden, Luc Devroye and
Siamak Ravanbahksh for the invaluable challenges and assistance along the way,
which sincerely enriched my appreciation for mathematics. This thesis would look
very different without your influence.

I am extremely indebted to the wonderful community of free\footnote{As in
freedom, of course, rather than beer} open source software developers that
deserve tons of credit for the role they've played in improving my productivity.
In particular, I thank the Gentoo developers, especially those that have
directly assisted me numerous times on IRC, Matthew Johnson and all the
other Jax developers for fixing issues unbelievably quickly, and Linus for obvious
reasons.

Lastly, I thank my family and friends for their encouragement over the years. In
particular, I am very grateful to my parents, Casey, and my grandparents for
their never-ending support.
